Network Configuration Synthesis with Abstract Topologies

### 2016-11-12 paper/abstraction/paper.tex
more edits

Based on these inputs, the \sysname compiler generates one template per role. It
guarantees that, when the templates are instantiated to configurations for {\em
any} concrete topology that complies with the abstract topology, the result
network will correctly implement the routing policy and meet the fault-tolerance
requirements. We show how fault-tolerance analysis can be done over abstract
topology itself using \todo{say something cool about about the inference rules
and such}.

@@

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, the contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item Design of new topology abstractions for network programming
based on graph homomorphisms and connectivity invariants.
\item Design of two new algorithms for analyzing routing policy over
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance
properties of network policy.  Both algorithms provide guarantees for any
concrete network satisfying the abstract connectivity invariants.
\item Design of template generation strategies that support network
evolution.
\item Implementation and evaluation of a \sysname compiler that generates
configuration templates and is up to two orders of magnitude faster
than its closest competitor on large networks.
\end{itemize}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/example3}
  \caption{An abstraction for the datacenters in Figure~\ref{fig:example}.}
  \label{fig:example3}
  \vspace{-1em}
\end{figure}

The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs. On its own, this abstraction loses a lot of information about the underlying structure of the concrete network, making it difficult to reason precisely about fault-tolerance.
%which would hinder a useful analysis of its fault-tolerance.
For example, with this abstraction any spine router in a concrete network may or may not be connected to any aggregator router.
%
In order to refine the abstraction to capture more precisely the kinds of concrete networks of interest,
we introduce several additional concepts. The first is the idea of a topology hierarchy captured by P and Q, which say that nodes in the ToR and aggregator roles are grouped into a number of smaller pods.
The second is a notion of node and edge \emph{multiplicities}. Each edge (and node) is labelled
with a symbolic variable (\EG, $e1, e2, \ldots$) that denotes a constraint on the number of edges (and nodes) that may appear in any given concrete network. Operators can capture concrete network invariants by adding constraints on the symbolic variables using logical formula.

@@

Next we can capture constraint (4) that traffic for local prefixes must stay local to the datacenter:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local constraint applies to any traffic destined for a local prefix described by the $\CD{\$LP}$ prefix template variable and adds the constraint that the traffic must follow a path that matches an internal location (\IE, a router inside the datacenter) at each hop of the path.
%
The constraint to prevent "valleys" (5) is implemented as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each datacenter level.
%
Constraint (7) to prevent transit traffic between neighbors is expressed in \sysname as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ to be either $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters the datacenter through a peer and exits through a peer. Finally, we can combine these constraints together as follows:

@@

\begin{defn}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete $PG = (G',\mathit{start},P)$ and abstract $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}

@@

If we determine that each router has a total order with respect to the simulation relation ($\leq$) for its corresponding PG nodes, then we can simply sort the nodes and prefer advertisements from peers of node $m$ over those of node $m'$ whenever $m' \leq m$. This condition ensures that BGP devices will converge to the most preferred paths in the network even when failures occur.
%
For example, in the concrete PG the inequality $(S1,0,1) \leq (S1,1,0)$ holds since nodes on the left side of the PG can always match transitions made on the right hand side with respect to the $\geq_{rank}$ relation. This relation does not hold the other way around since $(S1,0,1) \ngeq_{rank} (S1,1,0)$. Therefore, advertisements received at $S1$ in state $(1,0)$ must be preferred to those received at $S1$ in state $(0,1)$.
%
Notice that in the abstract PG $(S,0,1) \leq (S,1,0)$ also holds. This leads us the the following observation:

@@

This gives us reason to believe that we can perform compilation over the abstract PG directly and ensure that the inferred device-level preferences will be correct when instantiating the template with a concrete network.

@@

The \sysname compiler is implemented in approximately 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate configurations for open-source Quagga routers by default. The fault-tolerance analysis uses $\nu$Z~\cite{z3opt} to test validity and to minimize variables subject to the annotated constraints. Since the analysis typically calls the SMT solver many times with relatively simple optimization problems, we use a short timeout of 200ms.

>>>>

Based on these inputs, the \sysname compiler generates one template per role. It
guarantees that, when the templates are instantiated to configurations for {\em
any} concrete topology that complies with the abstract topology, the result
network will correctly implement the routing policy and meet the fault-tolerance
requirements. We show how fault-tolerance analysis can be done over the abstract
topology itself by inferring the worst-case number of edge-disjoint paths in some concrete topology using an optimizing SMT solver.

@@

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, the contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item Design of new topology abstractions for network programming
based on graph homomorphisms and connectivity invariants.
\item Design of two new algorithms for analyzing routing policy over
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance
properties of network policy.  Both algorithms provide guarantees for any
concrete network satisfying the abstract connectivity invariants.
\item Design of a template generation strategy that supports network
evolution.
\item Implementation and evaluation of a \sysname compiler that generates
configuration templates and is up to two orders of magnitude faster
than its closest competitor on large networks.
\end{itemize}

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/example3}
  \vspace{-2em}
  \caption{An abstraction for the datacenters in Figure~\ref{fig:example}.}
  \label{fig:example3}
  \vspace{-1em}
\end{figure}

The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs. On its own, this abstraction loses a lot of information about the underlying structure of the concrete network, making it difficult to reason precisely about fault-tolerance.
%which would hinder a useful analysis of its fault-tolerance.
For example, with this abstraction any spine router in a concrete network may or may not be connected to any aggregator router.
%
In order to refine the abstraction to capture more precisely the kinds of concrete networks of interest,
we introduce several additional concepts. The first is the idea of a topology hierarchy captured by P and Q, which indicate that nodes in the ToR and aggregator roles are grouped into a number of smaller pods.
The second is a notion of node and edge \emph{multiplicities}. Each edge (and node) is labelled
with a symbolic variable (\EG, $e1, e2, \ldots$) that denotes a constraint on the number of edges (and nodes) that may appear in any given concrete network. Operators can capture concrete network invariants by adding constraints on the symbolic variables using logical formula.

@@

Next we can capture constraint (4) that traffic for local prefixes must stay local to the datacenter:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local constraint applies to any traffic destined for a local prefix described by the $\CD{\$LP}$ prefix template variable and adds the constraint that the traffic must follow a path that matches an internal location (\IE, in the datacenter) at each hop of the path.
%
The constraint to prevent "valleys" (5) is written:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each datacenter level.
%
Constraint (7) to prevent transit traffic between neighbors is expressed in \sysname as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ to be either $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters the datacenter through a peer and exits through a peer. Finally, we can combine these constraints together as follows:

@@

\begin{defn}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete $PG = (G',\mathit{start},P)$ and abstract $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$ as follows:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}

@@

If we determine that each router has a total order with respect to the simulation relation ($\leq$) for its corresponding PG nodes, then we can simply sort the nodes and prefer advertisements from peers of node $m$ over those of node $m'$ whenever $m' \leq m$. This condition ensures that BGP devices will converge to the most preferred paths in the network even when failures occur.
%
For example, in the concrete PG the inequality $(S1,0,1) \leq (S1,1,0)$ holds since nodes on the left side of the PG can always match transitions made on the right hand side with respect to the $\geq_{rank}$ relation. This relation does not hold the other way around since $(S1,0,1) \ngeq_{rank} (S1,1,0)$. Therefore, advertisements received at $S1$ in state $(1,0)$ must be preferred to those received at $S1$ in state $(0,1)$.
%
Notice that in the abstract PG $(S,0,1) \leq (S,1,0)$ also holds.

@@

Lemma~\ref{lem:preference} gives us reason to believe that we can perform compilation over the abstract PG directly and ensure that the inferred device-level preferences will be correct when instantiating the template with a concrete network.

@@

The \sysname compiler is implemented in around 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate configurations for open-source Quagga~\cite{quagga} routers by default. The fault-tolerance analysis uses $\nu$Z~\cite{z3opt} to test validity and minimize variables subject to the topology constraints. Since the analysis typically calls the SMT solver many times with relatively small optimization problems, we use a timeout of 200ms.
